### 2022-09-08 manu-draft.qmd
proofing intro, methods, updating workflow graphic

---
format: 
  docx:
    reference-doc: my_styles.docx
title: "Addressing climate change and development pressures in an urban estuary through habitat restoration planning"
shorttitle: "Tampa Bay watershed habitat restoration planning"
bibliography: refs.bib
csl: frontiers-in-ecology-and-evolution.csl
author: 
  - Marcus W Beck:
      email: mbeck@tbep.org
      institute: [TBEP]
      correspondence: true
  - Douglas E Robison:
      email: DRobison@esassoc.com 
      institute: [ESA]
      correspondence: false
  - Gary E Raulerson:
      email: gmangrove@gmail.com
      institute: [TBEP]
      correspondence: false
  - Maya C Burke:
      email: mburke@tbep.org
      institute: [TBEP]
      corrspondence: false
  - Justin Saarinen:
      email: justinsaarinen@gmail.com 
      institute: [ESA]
      correspondence: false
  - Christine Sciarrino:
      email: CSciarrino@esassoc.com 
      institute: [ESA]
      correspondence: false
  - Edward T Sherwood:
      email: esherwood@tbep.org 
      institute: [TBEP]
      correspondence: false
  - David A Tomasko:
      email: dave@sarasotabay.org
      institute: [SBEP]
      correspondence: false
institute:
  - TBEP: Tampa Bay Estuary Program, St. Petersburg, Florida 33701 USA
  - ESA: Environmental Science Associates, Tampa, Florida 33609 USA
  - SBEP: Sarasota Bay Estuary Program, Sarasota, Florida 34236 USA

filters: 
  - templates/scholarly-metadata.lua
  - templates/author-info-blocks.lua

execute: 
  echo: false
  warning: false
  ft.keepnext: false
  ft.align: left
---

```{r, results = 'hide'}
library(knitr)
library(here)
library(flextable)
library(dplyr)

load(file = here('tabs/subttab.RData'))
load(file = here('tabs/acretab.RData'))
load(file = here('tabs/currtab.RData'))
load(file = here('tabs/targtab.RData'))
load(file = here('data/areas.RData'))
```

## Abstract

Native habitats in Florida face dual pressures at the land-sea interface from urban development and sea-level rise. To address these pressures, restoration practitioners require robust tools that identify reasonable goals given historical land use trends, current status of native habitats, and anticipated future impacts from coastal stressors. A target-setting approach for native habitats was created for the Tampa Bay watershed that identifies current restoration opportunities and establishes short-term (2030) targets and long-term (2050) goals. The approach was informed through a three-decade habitat change analysis and over forty years of habitat restoration projects in the region. Restoration goals were defined based on what is possible today and the projected needs for the future, rather than attempting to replicate past ecological conditions. The new approach also accounts for the expected impacts of sea-level rise, climate change, and watershed development - stressors which are pervasive in Florida. The resulting habitat goals are spatially explicit with maps that identify remaining restoration opportunities, while also providing an approach for the entire watershed that targets subtidal, intertidal, and uplands. This approach also supports coastal planning decisions that need to address competing interests and could be applied in other coastal settings where sustainable urbanization practices need to co-exist with natural environments. Methods for repeatable analyses are also available using an open source workflow to update progress over time and for adoption by others.

## Introduction

The health of estuarine systems and coastal habitats is tightly linked to land use and management of the watershed [@Yoskowitz15]. Coastal habitats provide multiple ecosystem services, including wildlife shelter and migratory corridors [@Yoskowitz15], fisheries production [@Houde93], water quality improvement [@Kushlan90; @Sprandel00; @AvilaGarcia20], erosion and flood attenuation [@Calil15; @Menendez18], carbon sequestration [@Dontis20] and recreation [@Chung18]. Pervasive stressors can negatively impact the services provided by these habitats. Anthropogenic land development can remove or substantially modify existing habitats, and can alter hydrology of nearby streams and rivers [@Theobald97; @Sim14]. Habitat changes in response to climate change include landward migration of mangroves into salt marshes, upstream migration of salt marshes within tidal tributaries, and upland forest migration [@Brinson95; @Vogelmann12; @Cavanaugh19]. Management of coastal habitats requires an understanding of how stressors impact these resources, while balancing competing societal uses.

The combined effects of land development and climate change are especially problematic for prioritizing habitat restoration activities in coastal environments. Landward migration of critical habitats in response to sea-level rise may not be possible due to anthropogenic barriers in the watershed. Sea-level rise can occur quicker than landward migration of salt marshes and the upland slope may already be lost to urban development and hardening [@Titus09]. The coastal squeeze of development and climate change diminishes the available space for future restoration in urban estuaries, thereby impacting ecosystem services provided by these habitats and the wildlife they support [@Enwright15]. Given projected habitat losses and the limited resources available, appropriate and realistic sites for restoration need to be identified that account for future stressors and past trends.

Past approaches for guiding restoration planning have been successfully used in other contexts, but they do not fully balance competing needs among public and private sectors, nor do they fully account for anticipated effects of multiple stressors. For example, an integrated watershed approach [@EPA96] has been utilized since the early 1990s to diagnose and manage water quantity and quality problems by addressing issues within hydrologically-defined geographic areas. Additionally, the habitat mosaic approach [@Henningsen05] of including multiple habitat types within restoration projects has been recognized as an effective means of allowing ecosystem state changes in response to different environmental pressures [@Duarte09; @Palmer09]. Adaptive management [@Holling78; @Gregory06] components have also been used to address challenges of sea-level rise, climate change, and development stressors, including monitoring to identify critical restoration decision points and needed intervention with contingency plans. Elements of each of these approaches could be combined to create a more holistic approach to guide restoration and conservation activities for coastal habitats in urban settings.

The Tampa Bay watershed (Florida, USA) is a valuable case study for developing a habitat restoration plan that addresses pervasive coastal stressors. Compared to other estuaries, the ratio of watershed to estuary area is small and the area is heavily developed with 42% of land use classified as urban and suburban residential [@SWFWMD18]. A retrospective approach to setting habitat protection and restoration targets in Tampa Bay was previously used [@Lewis96; @Robison10; @Cicchetti11; @Russell15]. Priority was given to restoration activities focused on habitat types that were important for a suite of estuarine faunal guilds disproportionately lost or degraded compared to a benchmark period. Criticisms of this approach included lack of consideration for future sea-level rise and other climate change factors [@Yoskowitz15], use of expanded and different habitats outside the Tampa Bay watershed [@Robison10], lack of attention to upland or freshwater wetland habitats, and little recognition of land development trends or actual available space for restoration efforts. These challenges are shared by restoration practitioners in other coastal environments and an approach that accommodates these challenges for planning would be highly transferable.

This paper describes an approach for habitat restoration and conservation planning that addresses the above challenges by considering the whole watershed, addressing historical changes, focusing on trajectories that have occurred during contemporary time periods, and considering both current and future stressors -- particularly land development and sea-level rise. Current and historical data are available for most Tampa Bay habitats, representing a time period when federal, state and local regulations were in effect and regional impacts from climate change have been documented [@Raabe12; @Cavanaugh14]. The approach establishes a framework that can guide both watershed-level habitat planning and site-level restoration activities and incorporates applicable elements of other habitat restoration paradigms discussed above [@Palmer09]. The general approach includes 1) designation of habitat types by strata relative to the aquatic-terrestrial gradient, 2) quantification of historical trends by habitat types to identify appropriate future targets in coverage, and 3) identification of opportunity areas that could be used by practitioners to achieve restoration goals based on habitat type and past trajectories. The outcomes are spatially-specific by providing maps to identity opportunity areas and reproducible using an open science workflow [@Lowndes17] that allows regular updates as new data become available.

## Methods

### Study area

Tampa Bay is a large open water estuary (surface area approximately 983 km^2^) on the west-central coast of Florida (@fig-lulcmap). The watershed covers approximately 5,872 km^2^, for a total combined area of approximately 6,855 km^2^. The climate is subtropical and within the 2020 ecotone for mangrove and salt marsh habitats. Naive habitats in the watershed include pine flatwoods, forested freshwater wetlands and non-forested vegetated wetlands. The watershed is heavily developed with an estimated population of 3.3 million people in the four major counties [@Rayer20]*.* Numerous anthropogenic changes have altered the natural habits of Tampa Bay, including direct removal of habitat (e.g., dredge and fill of bay bottom, mining activities), alteration of hydrology, and destruction and fragmentation of habitat from development.

### Habitats of Tampa Bay

The major habitat types of Tampa Bay were stratified by tidal influence and location in the watershed to define broad categories for restoration planning. Subtidal habitats included those that are submerged all or most of the time, intertidal habitats included emergent tidal wetlands that are submerged during high tides but exposed during low tides, and supratidal habitats included those that occur above the high tide line (i.e., the remainder of the watershed).

Subtidal habitats included hard bottom [@Jaap90; @Ash05], artificial reefs [@Dupont08], tidal flats [@Eisma98; @Moore68]*,* seagrasses [@Heck03; @Sherwood17], and oyster reefs [@Coen07; @zuErmgassen13]. Intertidal habitats (or emergent tidal wetlands) included mangroves [@Odum90], salt marshes [@Raabe12; @Comeaux12], salt barrens [@Berntness85; @Hsieh04]*,* tidal tributaries [@Sherwood08; @Wessel22], and living shorelines [@Smith18; @NOAA15; @RAE15]. Supratidal habitats included non-developed uplands [@Meyers90], freshwater forested wetlands [@Conner07], and freshwater non-forested wetlands [@Kushlan90]*.* Uplands were further sub-divided into coastal and non-coastal uplands based on location relative to the 5-foot contour that extended landward from the bay shoreline (described below). This 5-foot contour is an area of intense urban development and is expected to be impacted by sea-level rise based on current estimates. Since 1946, the St. Petersburg tidal gauge (NOAA gauge 872650) has documented a nearly 20 cm increase in mean tidal height to present day. Projections to the year 2100 suggest sea levels can increase between 58 and 259 cm in the region [@Burke19].

### Approach

Coverage targets for habitat types and opportunity areas for restoration were identified by integrating multiple datasets available for the region. First, habitat status and historical trends were quantified using land use/land cover and subtidal datasets to understand relative changes that have occurred over time. Second, historic habitat restoration efforts conducted in the watershed were synthesized as a guide for understanding a feasible level of effort for future projects that could be conducted by restoration practitioners. The first two steps were used to identify short-term (2030) targets and long-term (2050) goals for coverage (hectares) of native habitats. Finally, remaining opportunities for restoration were spatially identified by combining current coverages with existing or proposed protected areas and areas anticipated to be impacted by sea-level rise. As such, the approach identifies goals and targets and provides spatially explicit information that identifies where restoration practitioners should pursue projects to meet these goals and targets.

### Habitat status and trends

For the majority of subtidal, intertidal and supratidal habitats, coverages were quantified from two routine spatial assessment programs conducted by the Southwest Florida Water Management District (SWFWMD). For subtidal habitats, 2018 data were used to estimate current coverage of seagrasses, tidal flats, and oysters [@SWFWMD19]. Historical datasets for subtidal habitats began in 1988 with updates occurring on an approximate bi-annual basis [@Sherwood17; @Tomasko20]*.* Oyster bed coverage has been routinely estimated in these data products beginning in 2014.

Current intertidal and supratidal habitat coverages were estimated using the 2017 SWFWMD Land Use Land Cover (LULC) maps [@SWFWMD18]. Land use and cover types (natural and developed) are classified following the Florida Land Use Cover and Forms Classification System [FLUCCS, @FDOT99; @SWFWMD14]*.* Mangroves, salt barrens, and salt marshes were reported individually. While the photointerpretation of specific freshwater wetland types is often difficult, forested wetlands and non-forested wetlands can be distinguished with these data. Therefore, all applicable FLUCCS codes representing natural freshwater wetlands were combined for these classifications. Native upland habitats were also combined in one classification. Historical estimates for all intertidal and supratidal habitats were also quantified starting with the earliest database in 1990 and occurring every two to three years until the current estimate in 2017.

To address data gaps for habitats not included in the routine SWFWMD datasets, results from special studies were compiled to obtain current estimates. These included hard bottom subtidal habitats, artificial reefs, tidal creeks, and living shorelines.  No information on historical trends is available for these habitats.

Finally, a habitat coverage change analysis between the terminal years of data (1988 to 2018 for subtidal, 1990 to 2017 for intertidal and supratidal) was conducted to understand how habitats were changing between types. This required an intersection of the data layers to quantify if habitat types were unchanged or changed for any given location and identifying the type of change (e.g., seagrass to tidal flats). The results were summarized as Alluvial diagrams showing relative proportions of habitat change by type and between years [@Allaire17].

### Restoration and enhancement projects

Restoration projects conducted over the past 40 years were quantified for each of the major habitat types to inform expectations for setting short-term targets and long-term goals. Restoration was defined as any activity that involved earthwork to reshape the land or the addition of structural elements (e.g., rock). Enhancement was defined as any activity not including earthwork that improved the environment (e.g., planting native vegetation, invasive species or debris removal, prescribed burns, etc.). Data were gathered from the SWFWMD Surface Water Improvement and Management Program, Federal Government Performance and Results Act reporting, the Tampa Bay Water Atlas (<https://www.tampabay.wateratlas.usf.edu/>), Tampa Bay Watch, and the Technical Advisory Committee of the Tampa Bay Estuary Program. The collected data included project name, year, description, size (area or length), and location (latitude and longitude). Data gaps were supplemented by archival research, site visits, contacting entities, and expert knowledge from local professionals.

### Opportunity areas and restoration potential

Spatially-explicit estimates of the opportunity areas and their restoration potential in the Tampa Bay watershed were obtained using a spatial analysis shown in @fig-gisworkflow. The two main processes included 1) binning existing datasets into relevant categories and 2) overlaying multiple datasets to identify opportunities. Opportunity areas were defined as locations where habitat protection and restoration activities are possible and where they could occur to attain the targets and goals described above. Identifying opportunity areas is necessary to quantify the restoration potential for a particular habitat type, which is a measure of what is actually possible given expected land use change and sea-level rise.

The Land Use Land Cover 2017 dataset from SWFWMD was used for binning existing coverages into the relevant habitat types in the intertidal and subtidal strata. All FLUCCS classification codes were placed into one of three categories. First, native habitats were those that included the full range of natural plant communities and other habitats that are endemic to the watershed Second, restorable habitats included existing altered but non-hardened and pervious FLUCCS codes that could potentially support native habitats through restoration. Third, existing development included developed land FLUCCS codes that are hardened and impervious (e.g., structures and pavement) and not suitable for habitat restoration.

After binning, the native and restorable lands were overlaid with additional layers to identify 1) reservation native and reservation restorable areas, and 2) existing and proposed native and restorable areas. Collectively, each of these unique products are considered the opportunity areas in the Tampa Bay watershed. The reservation native and reservation restorable areas are native and restorable habitats, respectively, that occur in the 5-foot contour or coastal stratum. This area extends from the local Mean Lower Low Water (MLLW) elevation to an elevation 5 feet above Mean Sea Level and is highly vulnerable to sea-level rise [@Burke19]. Native habitats in this stratum were identified as those to be reserved, whereas restorable habitats were identified as those where tidal wetlands could be restored.

The existing and proposed native and restorable areas were those that occurred in public lands that are currently acquired or proposed for acquisition. To identify these areas, native and restorable lands were intersected with data created from the Florida Natural Areas Inventory and inventorying conservation and drainage easements data. This workflow created the existing conservation and proposed conservation layers in @fig-gisworkflow . Intersecting the native and restorable lands in these areas produced four unique opportunity areas: existing conservation native, proposed conservation native, existing conservation restorable, and proposed conservation native.

All opportunity areas identified as restorable included reservation restorable (in the coastal stratum), existing conservation restorable, and proposed conservation restorable. To identify discrete habitat types that could be the goal of future restoration projects, restorable lands in the coastal stratum and on existing conservation areas (reservation restorable and existing conservation restorable) were further grouped into their restoration potential. Proposed conservation areas were excluded from the analysis to provide a more confident assessment of restoration potential in areas that have already been acquired (i.e., existing conservation) or are immediately threatened by sea-level rise and/or coastal development (i.e., the coastal stratum). The restoration potential was defined using soil types, described as follows.

Compared to vegetation communities, soil characteristics typically change slowly (e.g., decades to centuries) in response to hydrologic impacts, unless physically disturbed [@Osland12; @Stockmann14]. Therefore, soil distributions can be used to estimate historical habitat distribution and restoration potential. A soils suitability layer was used for the Tampa Bay watershed [@Ries14] that classified soils as xeric, mesic, or hydric. The mesic and hydric categories were combined to represent wetland restoration potential and the xeric category was used to represent upland restoration potential. A distinction was made between tidal and freshwater wetland restoration potential by intersecting the mesic and hydric soils with the coastal stratum. Mesic or hydric soils that occur below the 5-foot contour were assigned a restoration potential for tidal wetlands, whereas mesic or hydric soils above the 5-foot contour were assigned a restoration potential for freshwater wetlands.

Two distinct mapping products were created from the above analysis. The first was an opportunities map that showed areas in the watershed identified as existing conservation native, existing conservation restorable, proposed conservation native, proposed conservation restorable, reservation native, and reservation restorable. The second was a map that identified the restorable lands (either existing or reservation) based on their restoration potential as coastal uplands, freshwater wetlands, native uplands, or tidal wetlands. All analyses were conducted using the R statistical programming language [@RCT22], specifically leveraging functions from the tidyverse package for data wrangling [@Wickham19] and the simple features (sf) package for geospatial analysis [@Pebesma18]. The workflows and data are provided in an open-access repository available on GitHub (<https://github.com/tbep-tech/hmpu-workflow>) [@Beck22].

## Results

### Habitat status and trends

Current estimates and trend information were available for seagrasses, tidal flats, and oyster bars (Table 1). Oyster bars were estimated at 67 ha in 2018 (Table 1), showing a 29% increase since mapping began in 2014. This increase may represent improved ground-truthing and photointerpretation of oyster bars from aerial photography. Tidal flats have generally increased from 1988 to the mid-2000s, followed by a decrease to present. The current estimate for tidal flats is 6,569 hectares, showing a 24% decline compared to the 1988 estimate of 8,700 hectares. Seagrasses have increased by 75% since 1988 to a current estimate of 6,986 ha. The change analysis comparing 1988 to 2018 for subtidal habitats (@fig-subtchgdatalluv) confirmed trends in Table 1 and showed that the seagrass increases were primarily associated with the colonization of non-vegetated areas of tidal flats, as well as unclassified areas of open water. Current estimates for subtidal habitats without historical trend information included 171 ha for natural hard bottom and 67 hectares for artificial reefs.

Total intertidal habitat (mangroves, salt barrens, and salt marshes) showed a 12% increase of 835 ha from 1990 to 2017, with a current estimate of 8,340 ha (Table 2). Mangroves increased by 15% to 6,276 ha, salt barrens increased by 7% to 203 ha, and salt marshes increased by 3% to 1,861 ha. Despite a net increase in salt marsh habitat, the change analysis showed that 153 ha were replaced by mangrove habitat (@fig-chgdatalluv). The current extent of tidal tributary length is 622 km (no trend information is available).

Trend assessments for supratidal habitats showed the effects of increasing land development and loss of restorable habitats in the Tampa Bay watershed (Table 2). Developed lands showed a 44% increase to 217,047 ha from 1990 to 2017, with a notable 30% decrease to 1,446 ha for coastal uplands, a 38% decrease to 57,836 ha for native uplands, and an 18% decrease to 189,512 ha for restorable lands. Non-forested freshwater wetlands showed a 24% increase to 27,358 ha, while forested freshwater wetlands decreased by 5% to 61,667 ha. The change analysis (@fig-chgdatalluv) showed that a majority of land conversion to developed came from restorable lands (21,292 ha) and native uplands (7,184 ha), with smaller proportions converted from forested freshwater wetlands (1,407 ha) and coastal uplands (193 ha). Habitats converted to restorable primarily included native uplands (8,304 ha), forested freshwater wetlands (1,700 ha), and developed (2,794 ha). The increase in non-forested freshwater wetlands was primarily from restorable lands (2,759 ha).

### Habitat restoration and enhancement

A total of 460 restoration projects were documented in Tampa Bay and its watershed between 1971 and 2019. These projects were divided among habitat types that included estuarine (n = 228), freshwater (n = 53), uplands (n = 119), and a mix of all three (n = 60). A total of 1,978 ha have been restored, whereas 12,930 ha and 42.8 km (as shoreline or tributaries) were enhanced. Forty partners were responsible for these projects, although some were from departments within the same agency. Eighty-nine living shoreline projects, seawall enhancements, and oyster reef installations were documented, with a total length of 18.2 km.  Although projects were documented for the whole period of record, few projects were completed prior to 1990.  From 1990 to 2010 and from 2010 to 2019, an annual mean of 68 ha/yr and 81 ha/yr of habitat was restored, respectively.  These means were used to define appropriate expectations for future restoration, described below.

### Summary of opportunity areas and restoration potential

Summaries of the current extent of each habitat type are shown in Table 3 as representations of the opportunity areas and restoration potential. The extent of each habitat in existing conservation lands and proposed conservation lands is shown. Summaries of the restoration potential under existing and proposed conservation lands is also shown. The majority of restoration opportunities on existing conservation lands are for native uplands and freshwater wetlands. Less opportunities exist for intertidal wetlands (mangrove forests, salt barrens, and salt marshes). These summaries are also shown spatially in @fig-oppmap and @fig-restmap.

The map of the remaining opportunity areas provided an areal estimate of where practitioners could target future restoration projects (@fig-oppmap). Native lands currently protected, proposed for protection, or in the coastal stratum (existing conservation native, proposed conservation native, reservation native) totaled `r areas$native$areaha` (`r areas$native$areaper` of the watershed above MLLW). Similarly, restorable lands currently protected, proposed for protection, or in the coastal stratum (existing conservation restorable, proposed conservation restorable, reservation restorable) totalled `r areas$restorable$areaha` (`r areas$restorable$areaper` of the watershed). Understandably, most of the native and restorable lands occurred in undeveloped areas in the north and southwest. Existing conservation lands (existing conservation native, existing conservation restorable) totalled `r areas$existing$areaha` (`r areas$existing$areaper` of the watershed) and proposed conservation lands (proposed conservation native, proposed conservation restorable) totalled `r areas$proposed$areaha` (`r areas$proposed$areaper` of the watershed). Reservation areas in the coastal stratum (reservation native, reservation restorable) totalled `r areas$reservation$areaha` (`r areas$reservation$areaper` of the watershed).

Combining the restorable lands on existing conservation areas and in the coastal stratum with soils data provided a spatial summary of the restoration potential grouped by habitat type (@fig-restmap). A total of `r areas$potentialall$areaha` (`r areas$potentialall$areaper` of the watershed) were identified as potentially restorable as coastal uplands (`r areas$potential[['Coastal Uplands']]$areaha`, `r areas$potential[['Coastal Uplands']]$areaper` of the watershed), freshwater wetlands (`r areas$potential[['Freshwater Wetlands']]$areaha`, `r areas$potential[['Freshwater Wetlands']]$areaper` of the watershed), native uplands (`r areas$potential[['Native Uplands']]$areaha`, `r areas$potential[['Native Uplands']]$areaper` of the watershed), or tidal wetlands (`r areas$potential[['Tidal Wetlands']]$areaha`, `r areas$potential[['Tidal Wetlands']]$areaper` of the watershed).

### Establishment of targets and goals

Identifying short-term (2030) targets and long-term (2050) goals for the restoration extent of native habitats in Tampa Bay was informed by the assessment of current extents, past trends, and relative effort for past restoration and enhancement projects. Table 4 shows the targets and goals identified through this analysis and the associated rationale. For example, the targets and goals are established based on the current extent and the restoration potential. If restoration potential exists and coverage restored from past projects suggests a reasonable level of effort, the targets and goals reflect the current extent plus the restoration opportunity. Conversely, other habitats with no identified restoration opportunity, or current extents are sufficient (e.g., mangrove forests), were assigned targets and goals similar to the current extent, i.e., these habitats should be protected and restoration is not recommended. Implicit in the targets and the goals is recurring re-assessment over time to evaluate progress and adjust expectations as appropriate.

## Discussion

Priorities for habitat restoration should be informed by current assessments and what is possible to achieve in the future. These priorities are necessary given anticipated impacts of land development and climate change, while also considering competing societal interests for use of the environment and limited resources for land acquisition and restoration. Our approach balances these tradeoffs by providing a systematic workflow to identify targets and goals that is informed by current extent, past trends, and realistic effort from past projects. Further, locations are identified where these targets and goals could be achieved based on existing opportunities for restorable habitats, including areas anticipated to be impacted by coastal stressors (i.e., sea level rise and land development). This approach departs from previous restoration paradigms by identifying what is possible rather than attempting to recreate historical baselines. Methods are also provided using open source tools [@Beck22] that 1) allow for the most current datasets to be synthesized to update progress on achieving targets and goals, and 2) can be used in other locations with similar needs for identifying restoration priorities.

### Habitat trends

Identifying appropriate targets and goals would not have been possible without a detailed assessment of current extent and past trends of native habitats in Tampa Bay and its watershed. The most notable trends included 1) the increase of seagrasses by 75%, 2) the increase of emergent tidal wetlands (12%) and freshwater wetlands (24%), and 3) the loss of native uplands (38%).

Seagrass recovery in Tampa Bay is a well-known success story that demonstrated how public-private partnerships can effectively reduce total nitrogen loads into Tampa Bay [@Greening14; @Sherwood17] . The nutrient reductions, primarily from point-source controls and advanced wastewater treatment, contributed to improvements in water quality and light environments that were favorable for seagrass growth. However, the most recent coverage estimate in 2020 showed a seagrass loss of 18% since the peak in 2016, falling below the target defined herein. These data were unavailable at the time the habitat restoration workflow was initially developed and trends informed by the new restoration paradigm have prompted bay managers to assess barriers in achieving the seagrass restoration goal. In particular, much of the seagrass losses have occurred in Old Tampa Bay, where recurring algal blooms of *Pyrodinium bahamense* have contributed to water quality decline [@Lopez19] . Ongoing research to understand mechanisms for mitigating these blooms will be critical for restoring seagrass in Tampa Bay.

Emergent tidal and freshwater wetlands in Tampa Bay have also experienced dramatic changes over the last three decades. Dual pressures from sea-level rise and climate change have impacted tidal wetlands, such that mangrove forests are outcompeting salt marshes and salt barrens for available niche space. Mangrove expansion as a result of climate change has been observed throughout the Gulf of Mexico [@Comeaux12]. As such, the identified targets and goals for mangroves indicate protection of these habitats, without the need for additional restoration. However, mangroves will continue to expand in the intertidal zone contributing to additional losses of salt marshes and salt barrens. The reservation areas identified in @fig-oppmap represent critical remaining areas in the intertidal zone that could be protected to prevent additional losses of tidal wetlands. Likewise, gains in non-forested freshwater wetlands are a reflection of 1) the effectiveness of state and federal wetland regulatory programs, and 2) the cumulative gains from publicly-funded habitat restoration projects. A similar approach could be used for salt marshes or salt barrens.

The decrease in native uplands is the result of continued development in the Tampa Bay watershed (@fig-chgdatalluv) and lack of regulatory protection of these habitats. Attaining the target and goal will require restoration of upland habitats on existing conservation lands (i.e., restoration potential in @fig-restmap) and new conservation lands to offset the continued loss of these habitats to development.  The long-term conversion of restorable lands to developed (@fig-chgdatalluv) presents additional challenges for restoration of native uplands.  Amendments to existing planning, zoning, and land development policies will be needed to address these issues. While federal and state regulations related to listed species management provide some protection to rare habitats, such as scrub jay (*Aphelocoma coerulescens*) habitat, common and historically abundant native habitats are largely unprotected (e.g., pine flatwoods).  Reductions in native uplands will likely continue unless local governments improve regulatory protections, such as strengthening language within comprehensive plans and development ordinances.

### Achieving restoration targets and goals

Achieving the defined targets and goals will require diverse approaches for habitat restoration and management.  Focusing efforts on publicly-owned conservation lands is expected to have long-term benefits and will be most cost-effective given the level of restoration effort compared to habitats that have already been impacted by anthropogenic activities. As such, public acquisition of remaining critical lands (e.g., coastal uplands) is a high priority given current development trends in the Tampa Bay watershed.  Other restoration targets (e.g., salt marshes) will not be feasible without additional public acquisition or public-private partnerships as a mechanism for doing so [@Holl01; @Benson18]. Therefore, varied approaches to leverage resources for restoration are needed and could include pursuing traditional grants, matching funds from multiple partners, and volunteer initiatives for habitat restoration. Recent gains in tidal and freshwater wetlands are primarily due to publicly-funded habitat restoration projects, state and federal wetland regulatory programs, and to a lesser extent, regulatory mitigation. Restoration activities for habitats without similar regulatory frameworks  should pursue the options above to achieve the defined targets and goals.

Other restoration activities could be pursued for the opportunity areas identified herein.  Substantial opportunities exist for upland restoration on reclaimed mined lands within the watershed (@fig-lulcmap, @fig-oppmap). For estuarine habitats, opportunity areas could include dredged holes or spoil disposal areas, either for enhancing existing subtidal habitats or creating areas that could be colonized by seagrasses. Some opportunities also exist on developed lands primarily through enhancement projects, although these have not been explicitly identified in the products created herein. Examples include the construction of living shorelines in place of hardened seawalls, placement of submerged habitat modules (e.g., artificial oyster reef) along urban shorelines, and creation of backyard habitats. Tidal tributary restoration could also include removal of salinity barriers and filling of dredged channel sections.  Overall, restoration practitioners must consider several options and choose those that are most feasible given the available resources and likelihood of success.

Creative approaches may be required in areas impacted by sea-level rise if land acquisition is not possible. These approaches are necessary to accommodate future landward migration of tidal wetlands or the protection of coastal uplands, while also preventing or minimizing development activities that can inhibit landward habitat shifts. Coastal setbacks, buffers, or public easements are traditionally used to restrict development within a given distance from the shoreline. However, rolling easements may be an alternative approach whereby protected areas are allowed to "roll" landward with expected changes in sea level rise. Rolling easements could disincentivize more intense urban development of low-lying coastal uplands in less developed agricultural or recreational land uses. Landowners could maintain current economic uses with a rolling easement, while reserving such lands to accommodate landward habitat migration.

Finally, wetland impacts and associated compensatory mitigation projects authorized under wetland regulatory programs could serve as possible restoration mechanisms. Mitigation activities have historically been conducted independent of watershed-level planning and monitoring processes. This disconnect has contributed to fragmented implementation and inconsistent compliance monitoring of mitigation projects, including historically poor documentation of wetland losses and gains in the Tampa Bay watershed. However, if properly focused and coordinated, compensatory mitigation activities could significantly contribute to the attainment of restoration targets and goals for the region.

### Limitations of the approach

Identifying restoration priorities was data intensive and would not have been possible without the resources available for the region. The workflow for identifying priorities required detailed and spatially-explicit datasets specific to the Tampa Bay watershed. Long-term datasets describing land use and cover in the watershed and the extent of in-bay habitats were necessary to categorize current extent and past trends. Similarly, supporting datasets included those that described existing and proposed conservation areas, soils, past restoration activities, and relevant spatial boundaries (i.e., watershed and coastline). Many of these datasets are available outside of Tampa Bay, although temporal and spatial resolutions may limit application to other areas. Additionally, considerable effort was made in working with regional partners to identify and fill knowledge gaps for relevant habitat types. For example, inventories of hard bottom habitats, living shorelines, tidal tributaries, and artificial reefs were created through special studies or were available only as current estimates from regional entities. Tracking progress towards targets in goals for these habitats will require regular updates to these datasets that are not part of routine land use and cover maps.

An additional assumption of the workflow, particularly for tracking progress, is that projects reported by partners will ultimately manifest in map products. Specifically, restoration effort by regional partners is cataloged in the available restoration database, which not only depends on voluntary reporting, but also represents a source of information on restoration extent that is separate from land cover maps. An expectation is that acreage restored for a given habitat as reported by a partner will ultimately be shown as a change in land use and cover on regional maps. The temporal lag between an actual project and how it may be reflected in a GIS product is unknown, which may create a disconnect between the updates in achieving targets and goals as new data layers are released and the effort reported by partners. The spatial resolution of mapping products may also be insufficient to detect habitat changes as reported in the restoration database. For these reasons, projects reported by partners are currently summarized separately from the assessments above that depend on GIS layers. Additional work is needed to reconcile these datasets for more streamline reporting.

### Conclusion

The establishment of targets and goals that account for climate change, development trajectories, land availability, and past restoration effort expands the restoration opportunities to a more comprehensive list of habitats for the entire watershed. Land acquisition is critical for attaining the defined targets and goals and will also provide new opportunities for outdoor access to the broader community. Successful restoration is also contingent on engaging multiple partners, non-governmental organizations, and private citizens. The products created herein will guide these efforts for the next thirty years by providing a continuously updated assessment of where the opportunities exist and if targets and goals are expected to be met.  The Tampa Bay area is not unique in its challenges for protecting native habitats and the approach is readily transferable to other locations where restoration priorities are needed in response to pervasive coastal stressors.

## Figures

![Land use and land cover for the Tampa Bay watershed, Florida, USA. The watershed includes the natural hydrologic boundary with minor modifications to include partners working with the Tampa Bay Estuary Program.](figs/lulcmap.png){#fig-lulcmap fig-align="center"}

![Spatial analysis workflow used to identify opportunity areas and restoration potential in the Tampa Bay watershed. Workflows are divided into binning of land use/land cover categories into relevant habitat types and spatial overlay of datasets to identify the opportunity areas.](figs/gisworkflow.png){#fig-gisworkflow fig-align="center"}

![Change analysis of habitat categories in the subtidal strata of the Tampa Bay watershed. The left column shows relative areas in 1988 and the right column shows relative areas in 2018 for each habitat category. The grey lines show the proportional change of each habitat category between the years.](figs/subtchgdatalluv.png){#fig-subtchgdatalluv fig-align="center"}

![Change analysis of habitat categories in the intertidal and supratidal strata of the Tampa Bay watershed. The left column shows relative areas in 1990 and the right column shows relative areas in 2017 for each habitat category. The grey lines show the proportional change of each habitat category between the years.](figs/chgdatalluv.png){#fig-chgdatalluv fig-align="center"}

![Opportunity areas for habitat conservation and reservation in the Tampa Bay watershed. Green indicates existing conservation, blue indicates proposed conservation, and pink indicates reservations opportunities. Each category is also grouped into native and restorable habitats. The outline is the Tampa Bay watershed.](figs/oppmap.png){#fig-oppmap fig-align="center"}

![Habitat restoration potential in the Tampa Bay watershed. Areas are identified as those where habitat restoration could target the four identified categories as coastal uplands, freshwater wetlands, native uplands, or tidal wetlands. Categories are based on the opportunity areas and soil types. The outline is the Tampa Bay watershed.](figs/restmap.png){#fig-restmap fig-align="center"}

\newpage

## Tables

```{r}
#| tbl-cap: Change over time in hectares for subtidal habitats in Tampa Bay.  Columns show years with available data and the final two columns show the change and percent change from 1988 to 2018.  Oyster bars were not meaningfully quantified prior to 2014. 
subttab
```

\newpage

```{r}
#| tbl-cap: Change over time in hectares for intertidal and supratidal habitats in Tampa Bay.  Columns show years with available data and the final two columns show the change and percent change from 1990 to 2017.
acretab
```

\newpage

```{r}
#| tbl-cap: "Summary of habitat restoration opportunities in the Tampa Bay watershed.  Summaries are based on 2017 land use data, 2018 subtidal data, best estimates for habitat types not in existing GIS layers, and current extent of existing and proposed conservation lands. Proposed conservation lands are those identified for acquisition. Total restoration opportunity does not account for lands currently existing or proposed for conservation. N/A: not applicable, I/D: insufficient data."
currtab
```

\newpage

```{r}
#| tbl-cap: "Recommended 2030 targets and 2050 goals for habitat restoration and protection in the Tampa Bay watershed. Targets and goals are based on 2017 land use data, 2018 subtidal data, best estimates for habitat types not in existing GIS layers, and current extent of existing and proposed conservation lands. Total restoration opportunity does not account for lands currently existing or proposed for conservation. N/A: not applicable, I/D: insufficient data."
targtab
```

\newpage

## References