### 2022-12-30 supplement.qmd

---
format: 
  docx:
    reference-doc: my_styles.docx
    number-sections: true
title: "Supplement: Addressing climate change and development pressures in an urban estuary through habitat restoration planning"
shorttitle: "Tampa Bay watershed habitat restoration planning"
bibliography: refs.bib
csl: frontiers-in-ecology-and-evolution.csl
author: 
  - Marcus W Beck:
      email: mbeck@tbep.org
      institute: [TBEP]
      correspondence: true
  - Douglas E Robison:
      email: DRobison@esassoc.com 
      institute: [ESA]
      correspondence: false
  - Gary E Raulerson:
      email: gmangrove@gmail.com
      correspondence: false
  - Maya C Burke:
      email: mburke@tbep.org
      institute: [TBEP]
      correspondence: false
  - Justin Saarinen:
      email: justinsaarinen@gmail.com 
      institute: [ESA]
      correspondence: false
  - Christine Sciarrino:
      email: CSciarrino@esassoc.com 
      institute: [ESA]
      correspondence: false
  - Edward T Sherwood:
      email: esherwood@tbep.org 
      institute: [TBEP]
      correspondence: false
  - David A Tomasko:
      email: dave@sarasotabay.org
      institute: [SBEP]
      correspondence: false
institute:
  - TBEP: Tampa Bay Estuary Program, St. Petersburg, Florida 33701 USA
  - ESA: Environmental Science Associates, Tampa, Florida 33609 USA
  - SBEP: Sarasota Bay Estuary Program, Sarasota, Florida 34236 USA

filters: 
  - templates/scholarly-metadata.lua
  - templates/author-info-blocks.lua

execute: 
  echo: false
  warning: false
  ft.keepnext: false
  ft.align: left
---

```{r, results = 'hide'}
library(knitr)
library(here)
library(flextable)
library(ftExtra)
library(dplyr)
```

```{r}
#| tbl-cap: "Table S1: Data sources for habitat types grouped by subtidal, intertidal, and supratidal strata."

data(strata)

tab <- strata %>%
  mutate(
    Source = case_when(
      HMPU_TARGETS == 'Hard Bottom' ~ '@SWFWMD18;@CSA19',
      HMPU_TARGETS %in% c('Artificial Reefs', 'Living Shorelines') ~ '@Robison20',
      HMPU_TARGETS %in% c('Seagrasses', 'Oyster Bars', 'Tidal Flats') ~ '@SWFWMD19',
      HMPU_TARGETS %in% c('Mangrove Forests', 'Salt Barrens', 'Salt Marshes') ~ '@SWFWMD18', 
      HMPU_TARGETS == 'Tidal Tributaries' ~ '@Wessel22',
      HMPU_TARGETS %in% c('Coastal Uplands', 'Non-Forested Freshwater Wetlands', 
                          'Forested Freshwater Wetlands', 'Native Uplands') ~ '@SWFWMD18'
    ), 
    HMPU_TARGETS = as.character(HMPU_TARGETS)
  ) %>% 
  na.omit() %>% 
  arrange(Category, HMPU_TARGETS) %>% 
  select(
    Stratum = Category, 
    `Habitat Type` = HMPU_TARGETS, 
    Source
  )

thm <- function(x){
  flextable::align(x, align = 'center', part = 'header') %>%
    flextable::align(align = 'left', j = 1, part = 'all') %>%
    flextable::align(align = 'left', j = 2, part = 'body') %>%
    flextable::border_inner() %>%
    flextable::width(width = 0.7, j = 1) %>%
    flextable::width(width = 5.8 / (ncol_keys(.) - 1), j = 2:ncol_keys(.)) %>%
    flextable::padding(padding = 0, part = 'all') %>%
    flextable::fontsize(size = 8, part = 'all')
}

tab %>% 
  as_grouped_data('Stratum') %>% 
  flextable() %>%
  colformat_md() %>%
  thm
```

## References {.unnumbered}